---
title: "pm566_hw4"
author: "Yiping Li"
output: github_document
date: "`r Sys.Date()`" 
always_allow_html: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#HPC
##Q1:
Rewrite the following R functions to make them faster
```{r}
# Total row sums
fun1 <- function(mat) {
  n <- nrow(mat)
  ans <- double(n) 
  for (i in 1:n) {
    ans[i] <- sum(mat[i, ])
  }
  ans
}

fun1alt <- function(mat) {
  # YOUR CODE HERE, wk10 lab q3
  rowSums(mat)
}

# Cumulative sum by row
fun2 <- function(mat) {
  n <- nrow(mat)
  k <- ncol(mat)
  ans <- mat
  for (i in 1:n) {
    for (j in 2:k) {
      ans[i,j] <- mat[i, j] + ans[i, j - 1]
    }
  }
  ans
}

fun2alt <- function(mat) {
  # YOUR CODE HERE
  t(apply(mat, 1, cumsum))
}

# Use the data with this code
set.seed(2315)
dat <- matrix(rnorm(200 * 100), nrow = 200)

# Test for the first
library(microbenchmark)
microbenchmark::microbenchmark(
  fun1(dat),
  fun1alt(dat), check = "equivalent"
)

# Test for the second
microbenchmark::microbenchmark(
  fun2(dat),
  fun2alt(dat), check = "equivalent"
)

#get rid off unit = "relative" due to error
```

##Q2: 
Make things run faster with parallel computing. 
The following function allows simulating PI
```{r}
sim_pi <- function(n = 1000, i = NULL) {
  p <- matrix(runif(n*2), ncol = 2)
  mean(rowSums(p^2) < 1) * 4
}

# Here is an example of the run
set.seed(156)
sim_pi(1000) # 3.132
```

In order to get accurate estimates, we can run this function multiple times, with the following code:
```{r}
# This runs the simulation a 4,000 times, each with 10,000 points
set.seed(1231)
system.time({
  ans <- unlist(lapply(1:4000, sim_pi, n = 10000))
  print(mean(ans))
})
```

Rewrite the previous code using parLapply() to make it run faster. Make sure you set the seed using clusterSetRNGStream():
```{r}
library(parallel)
cl <- makePSOCKcluster(4)
clusterSetRNGStream(cl, 1231)
clusterExport(cl,varlist = c("sim_pi"), envir = environment())

# YOUR CODE HERE
system.time({
  # YOUR CODE HERE
  ans <- unlist(parLapply(cl, 1:4000, sim_pi, n = 10000))
  print(mean(ans))
  stopCluster(cl)
})
```


#SQL
Setup a temporary database by running the following chunk
```{r}
# install.packages(c("RSQLite", "DBI"))
library(RSQLite)
library(DBI)

# Initialize a temporary in memory database
con <- dbConnect(SQLite(), ":memory:")

# Download tables
film <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film.csv")
film_category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/film_category.csv")
category <- read.csv("https://raw.githubusercontent.com/ivanceras/sakila/master/csv-sakila-db/category.csv")

# Copy data.frames to database
dbWriteTable(con, "film", film)
dbWriteTable(con, "film_category", film_category)
dbWriteTable(con, "category", category)
```

When you write a new chunk, remember to replace the r with sql, connection=con. Some of these questions will reqruire you to use an inner join. Read more about them here https://www.w3schools.com/sql/sql_join_inner.asp

##Q1: 
How many many movies is there avaliable in each rating catagory. (#or COUNT(*) AS N)
```{sql, connection=con}
SELECT rating, COUNT (film_id) AS movie_count
FROM film
GROUP BY rating
ORDER BY movie_count
```

##Q2: 
What is the average replacement cost and rental rate for each rating category.
```{sql, connection=con}
SELECT rating, 
        AVG(replacement_cost) AS avg_replacementcost,
        AVG(rental_rate) AS avg_rentalrate
FROM film
GROUP BY rating
```

##Q3: 
Use table film_category together with film to find the how many films there are witth each category ID
```{sql, connection=con}
SELECT category_id, COUNT(category_id) AS film_count
FROM film_category AS fc INNER JOIN film AS f 
ON fc.film_id = f.film_id
GROUP BY category_id
```

##Q4: 
Incorporate table category into the answer to the previous question to find the name of the most popular category.
```{sql, connection=con}
SELECT fc.category_id, name, COUNT(fc.category_id) AS film_count
FROM film_category AS fc INNER JOIN film AS f 
ON fc.film_id = f.film_id
LEFT JOIN category AS c
ON fc.category_id = c.category_id
GROUP BY fc.category_id
```
